---
title: "Assignment_b1"
author: "STAT545"
date: "2024-10-24"
output: github_document
Roxygen: list(markdown = TRUE)
---

### Making and documenting the function:

```{r}
#' Count Unique Values by Group
#'
#' This function calculates the number of unique values in a specified column (`summ_var`) within each group defined by one or more grouping variables. It’s useful for understanding the distribution of unique values within groups.
#' 
#' @param data A data frame or tibble containing the data to be summarized. Chose `data` as it clearly indicates the primary data input.
#' @param summ_var The variable for which unique values are counted within each group. Named `summ_var` to clearly denote that this is the summarized variable in the count.
#' @param ... One or more grouping variables within the dataset. These variables define the grouping structure for summarizing `summ_var`. The `...` notation allows for flexible input of single or multiple variables.
#' @param .groups A character string indicating how to handle grouping in the result. Options are "drop" (default), "keep", "drop_last", or "rowwise". The `.groups` name follows `dplyr` syntax for familiarity.
#'
#' @return A tibble with one row per group and a column `unique_sum`, showing the count of unique values in `summ_var` for each group defined by `...`.
#' @export
#' 
#' 
#' 

count_unique <- function(data, summ_var, ..., .groups = "drop") {
  data %>%
    group_by(...) %>%
    summarize(unique_sum = n_distinct({{ summ_var }}[!is.na({{ summ_var }})]), .groups = .groups)
}

```

### -------------------------------------------------

```{r}
library(dplyr)
library(gapminder)
library(testthat)
library(palmerpenguins)
```

### Examples of use of this function:

**Example 1. Count unique countries by continent**

By counting the unique countries by continent, it gives us a sense of how many distinct countries are represented per continent.

```{r, example 1}
count_unique(gapminder, country, continent)

```

**Example 2. Count Unique Countries by Continent and Year**

By using both continent and year as grouping variables, we see the number of unique countries for each year within each continent.

```{r, example 2}
count_unique(gapminder, country, continent, year, .groups = "keep")
```

**Example 3: Deliberately causing an error by calling a non-existent variable**

In this example, \`non_existent_var\` does not exist in the \`gapminder\` dataset, so an error will be triggered when trying to summarize by non_existent_var. By setting \`error = TRUE\`, the document will render normally, and the error message will display in the output.

```{r, error = TRUE}
count_unique(gapminder, non_existent_var, continent)

```

### Testing this function:

**Testing vector with no NA’s.** There are no NA's in the species variable of the *palmerpenguins* dataset.

This test checks if the summary showing each island and the number of unique species observed on it are the same with my function and with using the *group_by %\>% summarize* function.

```{r, no NA}
sum(is.na(penguins$species))

test_that("count_unique function correctly counts unique species by island", {
  
  unique_species_count <- penguins %>%
    group_by(island) %>%
    summarize(unique_sum = n_distinct(species), .groups = "drop")

  expect_equal(count_unique(penguins, species, island), unique_species_count)

})
```

**Testing vector that has NA’s, also tests with a numerical variable instead of categorical.** There are 2 NA's in the numerical *flipper_length_mm* variable of the *palmerpenguins* dataset.

This test checks if the summary showing the number of unique flipper_length_mm observed on for each species are the same with my function and with using the *filter is.NA %\>% group_by %\>% summarize* function.

```{r, with NA, numerical}
sum(is.na(penguins$flipper_length_mm))

test_that("count_unique function handles missing values in `flipper_length_mm`", {
  
  expected_result <- penguins %>%
    filter(!is.na(flipper_length_mm)) %>% 
    group_by(species) %>%
    summarize(unique_sum = n_distinct(flipper_length_mm), .groups = "drop")
  
  expect_equal(count_unique(penguins, flipper_length_mm, species), expected_result)
})
```

**Test 3: Handling an empty dataset.** Filtering penguins to 0 row renders the dataset empty with no rows.

This test checks if an empty input dataset (resulting from a filter that yields no matching rows) will correctly return an output with zero rows.

```{r, empty dataset}
test_that("count_unique function handles vector of length 0 correctly", {

  empty_penguins <- penguins %>% filter(island == "Nonexistent")
  
  expect_equal(nrow(count_unique(empty_penguins, species, island)), 0) 
})
```
